Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations

### 2016-02-08 paper.tex
Getting all LU algorithms in there

\section{Preliminaries}
The Navier-Stokes equations used to describe turbulent flow in a doubly 
periodic channel can be written as
\begin{align}
 \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
 \nabla^2 \bm{u} - \nabla{p}, \notag \\
 \nabla \cdot \bm{u} &= 0, \label{eq:NS}
\end{align}
where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
and $t$ are position and time, $\bm{\mathcal{H}}(\bm{x}, t)$ represents the 
nonlinear convection and the mean pressure gradient (the driving force), $\nu$ 
is a constant dynamic viscosity and $p$ is the instantaneous pressure 
fluctuation normalized by a constant density. The 
computational domain is $\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$ with 
channel walls located at $x=\pm 1$ such that no-slip applies at $ \bm{u}(\pm 1, 
y, z, t) = 0$. The walls are spanning the $y-z$ plane and the equations are 
periodic in the $y$ and $z$ directions.

@@

Three-dimensional basis functions and function spaces that are periodic in $y$ 
and $z$ directions, can now be defined as
\begin{align}
  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
\end{align}
where $\imath=\sqrt{-1}$ and 
\begin{align}
K_N = \Big\{\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / &(l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
 &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\}. \label{eq:wavenumbermesh}
\end{align}
$\D{K}_N$ and $\N{K}_N$ are both equal to $K_N$ except from index $l$ that 
ranges from $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). The 
computational mesh has in real space been created using $N = (N_x, N_y, N_z)$ 
intervals, where the two periodic directions use uniform intervals. The 
computational mesh is given as
\begin{align}
  X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\} \label{eq:Xn}
\end{align}
where $x_i = h(i)$ represents either
\begin{equation}
 h(i) = \begin{cases}
   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Lobatto points}, \\
   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Chebyshev points}. \\
 \end{cases}
\end{equation}

@@

If the expansion coefficients $\hat{u}$  are known for the entire wavenumber 
mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_solx}) may be 
evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse 
transforms as
\begin{align}
u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n 
\underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l 
\hat{u}(l,m,n)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath \underline{m} 
y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , 
\quad \forall \, \bm{x} \in X_N, \notag \\
  u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
  (\N{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  
  \label{eq:ifft} 
\end{align}
Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represents inverse 
Fourier transforms along directions $y$ and $z$ respectively, and the inverse 
Shen transform $\N{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal 
$x$ direction. Note that the transforms are applied sequentially to the entire 
computational mesh. The inverse Shen transform may be computed using fast 
Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
\ref{eq:phiN}). The transforms are slightly different for the three bases and 
${\mathcal{S}}, \D{\mathcal{S}}$ and $\N{\mathcal{S}}$ are used to distinguish 
between them with obvious notation. Python implementations of forward and 
inverse Shen transforms and scalar products are given in the Appendix.

A three-dimensional scalar product in the weighted $[L^2_w(\Omega)]^3$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, \notag \\
   &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}\,dx_wdydz,
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction, and we have used $dx_w = dx/\sqrt{1-x^2}$. In this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, where quadrature is employed for integration. As such the scalar product corresponds to sums that may be computed using fast transforms
\begin{align}
 \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
 \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
 \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
 }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
  &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
  (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
\end{align}
Here $h\N{\mathcal{S}}$ denotes the complete three dimensional scalar product 
and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
$\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and 
$y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
\N{\phi}_l)_w$ is the Shen scalar product in the $x$-direction for the 
$\N{V}_N$ space. The weights, $w_i$, required for the Shen transforms are 
defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
       Gauss-Lobatto points},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
       Gauss-Chebyshev points},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.

The scalar product  $h\N{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a 
complete transform. To find a transformation between physical $u$ and spectral 
$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly and use the discrete 
orthogonality of the Fourier basis functions
\begin{align}
\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
\N{\psi}_{\bm{k}} \right>_w, \notag \\
           &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
           \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
           &= h \sum_{p=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t), 
           \label{eq:sst2}
\end{align}
where the scalar product $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_w = 
\sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
pentadiagonal mass matrix. Likewise, for the remaining two velocity components, 
the corresponding mass matrix, $\D{B}_{lq} = (\D{\phi}_q, \D{\phi}_l)_w$, is 
tridiagonal. The complete transformation is now obtained by setting Eq. 
(\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. Moving to 
matrix notation we get
\begin{align}
h\sum_{q=0}^{N_x-4} \N{B}_{lq}\, \hat{u}(q, {m}, {n}, t) &= 
h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
 \hat{u}_{\bm{k}}(t) &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
 \quad \forall \, \bm{k} \in \N{K}_N.
\end{align}
where $\N{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = 
\N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that since $\N{B}$ assembles to 
a pentadiagonal matrix the solution ($\N{B}^{-1}$) can be obtained very fast 
and 
the complete 
transformation thus requires a fast Chebyshev transform ($O(N_x \log N_x)$) and 
a fast linear algebraic solve ($O(N_x)$). Similar transforms $\mathcal{T}$ 
and $\D{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and 
(\ref{eq:phiD}), using mass matrices $B_{lq}=(T_q, T_l)_w$ (diagonal) and 
$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (tridiagonal) and scalar products 
$\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
\D{\phi}_l)_w$.

\section{Discretization of Navier Stokes equations}
\label{sec:discretizationNS}
The Navier Stokes equations (\ref{eq:NS}) are solved using a scheme 
proposed by Kim, Moin and Moser (1987). This scheme is developed by taking the 
Laplacian of the wall normal momentum equation and the curl of the momentum 
equation. Following elimination of the pressure, the equations to solve are 
\begin{align}
\frac{\partial}{\partial t} \nabla^2 u &= h_u + \nu \nabla^4 u, 
\label{eq:biharmonic} \\
\frac{\partial g}{\partial t} &= h_g + \nu \nabla^2 g, \label{eq:g} \\
f + \frac{\partial u}{\partial x} &= 0, \label{eq:f}
\end{align}
where
\begin{align}
f &= \frac{\partial v}{\partial y} + \frac{\partial w}{\partial z}, \\
g &= \frac{\partial w}{\partial y} - \frac{\partial v}{\partial z}, \\
h_u = -\frac{\partial}{\partial x} &\left( \frac{\partial 
\mathcal{H}_y}{\partial y} + \frac{\partial \mathcal{H}_z}{\partial z} \right) 
+ \left(\frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2} 
\right) \mathcal{H}_x ,
\\
h_g &= \frac{\partial \mathcal{H}_z}{\partial y} - \frac{\partial 
\mathcal{H}_y}{\partial z}.
\end{align}
The two remaining velocity components are computed from the definitions of $f$ 
and $g$. The biharmonic equation (\ref{eq:biharmonic}) is solved with boundary 
conditions are $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann condition follows 
from the continuity equation. The boundary conditions for Eq. (\ref{eq:g}) are 
$g(\pm 1) = 0$. The variational formulation becomes: Find ${u} \in 
\N{V}_N$, ${g} \in \D{V}_N$ and ${f} \in \D{V}_N$ such that
\begin{align}
	\frac{\partial }{\partial t} \left< \nabla^2 u, \N{\psi}\right>_w &= 
	\left<h_u, \N{\psi} \right>_w + \nu \left<\nabla^4u, \N{\psi}\right>_w 
	&\forall \N{\psi} \in \N{V}_N, \label{eq:u1} \\
	\frac{\partial}{\partial t}\left<g, \D{\psi}\right>_w &= \left<h_g, 
	\D{\psi}\right>_w + \nu 
	\left<\nabla^2 g, \D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N ,
	\label{eq:g1} \\
	\left<f, \D{\psi}\right>_w &= \left<\frac{\partial u}{\partial x}, 
	\D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N. \label{eq:f1}
\end{align}
The two remaining velocity components are computed by projection to the 
Dirichlet space $\D{V}_N$: Find 
${v}, {w}$ in $\D{V}_N \times \D{V}_N$ such that
\begin{align}
\left<f, \D{\psi}\right>_w &= \left<\frac{\partial v}{\partial y} + 
\frac{\partial w}{\partial z}, \D{\psi}\right>_w, \, &\forall \D{\psi} \in 
\D{V}_N \label{eq:f2} \\
\left<g, \D{\psi}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
\frac{\partial v}{\partial z}, \D{\psi}\right>_w, \, &\forall \D{\psi} \in 
\D{V}_N, \label{eq:g2}
\end{align}
which simplifies considerably because all the derivatives are in periodic 
directions. Written in spectral space Eqs. (\ref{eq:f2}) and (\ref{eq:g2}) 
become simply
\begin{align}
\hat{f}_{\bm{k}} &= \imath \underline{m}\, \hat{v}_{\bm{k}} + \imath 
\underline{n}\, \hat{w}_{\bm{k}} &\forall \bm{k} \in \D{K}_N^0, \label{eq:f3} \\
\hat{g}_{\bm{k}} &= \imath \underline{m}\, \hat{w}_{\bm{k}} - \imath 
\underline{n}\, \hat{v}_{\bm{k}} & \forall \bm{k} \in \D{K}_N^0, \label{eq:g3}
\end{align}
where $\D{K}_N^0$ is used to denote that these equations can be solved for all 
wavenumbers except $m=n=0$. For $m=n=0$ we solve instead the 
corresponding momentum equations in $y$ and $z$ directions:
\begin{align}
\frac{\partial }{\partial t} \D{B}_{lq}\hat{v}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) + \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) 
& \forall\, l \in [0, 1, 
\ldots, N_x-2], \label{eq:v0}\\
\frac{\partial }{\partial t} \D{B}_{lq}\hat{w}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_z(q, 0, 0, t) + \nu \D{A}_{lq} \hat{w}(q, 0, 0, t) 
& \forall\, l \in [0, 1, \ldots, N_x-2].\label{eq:w0}
\end{align}

>>>>

\section{Preliminaries}
The Navier-Stokes equations used to describe turbulent flow in a doubly 
periodic channel can be written in rotational form as
\begin{align}
 \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
 \nabla^2 \bm{u} - \nabla{P}, \notag \\
 \nabla \cdot \bm{u} &= 0, \label{eq:NS}
\end{align}
where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
and $t$ are position and time and $\bm{\mathcal{H}}(\bm{x}, t) = \bm{u}\times 
\bm{\omega}$ (where $\bm{\omega} = \nabla \times \bm{u}$)  represents the 
nonlinear convection.  The constant dynamic viscosity is denoted as $\nu$ and 
$P$ is a pressure modified to account for the kinetic energy, i.e., $P = p + 
\bm{u} \cdot \bm{u}/2$, where $p$ the instantaneous pressure 
normalized by a constant density. The computational domain is $\Omega=[-1, 
1]\times [0, L_y] \times [0, L_z]$ with channel walls located at $x=\pm 1$ such 
that no-slip applies at $ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning 
the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions 
with periodic lengths $L_y$ and $L_z$, respectively.

@@

Three-dimensional basis functions and function spaces that are periodic in $y$ 
and $z$ directions, can now be defined as
\begin{align}
  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
\end{align}
where $\imath=\sqrt{-1}$ and 
\begin{align}
K_N = \Big\{\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / &(l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
 &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\}. \label{eq:wavenumbermesh}
\end{align}
$\D{K}_N$ and $\N{K}_N$ are both equal to $K_N$ except from index $l$ that 
ranges $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). We use 
notation $K_N^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
dimensions of $K_N$.

The computational mesh has in real space been created using $N = (N_x, N_y, 
N_z)$ 
intervals, where the two periodic directions use uniform intervals. The 
computational mesh is given as
\begin{align}
  X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\} \label{eq:Xn}
\end{align}
where $x_i = h(i)$ represents either
\begin{equation}
 h(i) = \begin{cases}
   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Lobatto points}, \\
   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Chebyshev points}. \\
 \end{cases}
\end{equation}

@@

If the expansion coefficients $\hat{u}$  are known for the entire wavenumber 
mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_solx}) may be 
evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse 
transforms as
\begin{align}
u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n 
\underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l 
\hat{u}(l,m,n)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath \underline{m} 
y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , 
\quad \forall \, \bm{x} \in X_N, \notag \\
  u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
  (\N{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  
  \label{eq:ifft} 
\end{align}
Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represents inverse 
Fourier transforms along directions $y$ and $z$ respectively, and the inverse 
Shen transform $\N{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal 
$x$ direction. Note that the transforms are applied sequentially to the entire 
computational mesh. The inverse Shen transform may be computed using fast 
Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
\ref{eq:phiN}). The transforms are slightly different for the three bases and 
${\mathcal{S}}, \D{\mathcal{S}}$ and $\N{\mathcal{S}}$ are used to distinguish 
between them with obvious notation.

A three-dimensional scalar product in the weighted $[L^2_w(\Omega)]^3$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, \notag \\
   &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}\,dx_wdydz,
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction, and we have used $dx_w = dx/\sqrt{1-x^2}$. In this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, where quadrature is employed for integration. As such the scalar product corresponds to sums that may be computed using fast transforms
\begin{align}
 \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
 \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
 \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
 }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
  &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
  (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
\end{align}
Here $h\N{\mathcal{S}}$ denotes the complete three dimensional scalar product 
and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
$\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and 
$y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
\N{\phi}_l)_w$ is the Shen scalar product in the $x$-direction. The weights, 
$w_i$, required for the Shen transforms are 
defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
       Gauss-Lobatto points},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
       Gauss-Chebyshev points},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.

The scalar product  $h\N{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a 
complete transform. To find a transformation between physical $u$ and spectral 
$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly and use the discrete 
orthogonality of the Fourier basis functions
\begin{align}
\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
\N{\psi}_{\bm{k}} \right>_w, \notag \\
           &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
           \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
           &= h \sum_{p=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t), 
           \label{eq:sst2}
\end{align}
where $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_w = 
\sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
pentadiagonal mass matrix. The complete transformation is now obtained by 
setting Eq. (\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. 
Moving to matrix notation we get
\begin{align}
h\sum_{q=0}^{N_x-4} \N{B}_{lq}\, \hat{u}(q, {m}, {n}, t) &= 
h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
 \hat{u}_{\bm{k}}(t) &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
 \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:fstB}
\end{align}
where $\N{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = 
\N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that since $\N{B}$ assembles to 
a pentadiagonal matrix the solution ($\N{B}^{-1}$) can be obtained very fast 
and 
the complete 
transformation thus requires a fast Chebyshev transform ($O(N_x \log N_x)$) and 
a fast linear algebraic solve ($O(N_x)$). Similar transforms $\mathcal{T}$ 
and $\D{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and 
(\ref{eq:phiD}), using mass matrices $B_{lq}=(T_q, T_l)_w$ (diagonal) and 
$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (tridiagonal) and scalar products 
$\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
\D{\phi}_l)_w$.

\section{Discretization of Navier Stokes equations}
\label{sec:discretizationNS}
The Navier Stokes equations (\ref{eq:NS}) are solved using a scheme 
proposed by Kim, Moin and Moser (1987). This scheme is developed by taking the 
Laplacian of the wall normal momentum equation and the curl of the momentum 
equation. Following elimination of the pressure, the equations to solve are 
\begin{align}
\frac{\partial}{\partial t} \nabla^2 u &= h_u + \nu \nabla^4 u, 
\label{eq:biharmonic} \\
\frac{\partial g}{\partial t} &= h_g + \nu \nabla^2 g, \label{eq:g} \\
f + \frac{\partial u}{\partial x} &= 0, \label{eq:f}
\end{align}
where
\begin{align}
f &= \frac{\partial v}{\partial y} + \frac{\partial w}{\partial z}, \\
g &= \frac{\partial w}{\partial y} - \frac{\partial v}{\partial z}, \\
h_u = -\frac{\partial}{\partial x} &\left( \frac{\partial 
\mathcal{H}_y}{\partial y} + \frac{\partial \mathcal{H}_z}{\partial z} \right) 
+ \left(\frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2} 
\right) \mathcal{H}_x ,
\\
h_g &= \frac{\partial \mathcal{H}_z}{\partial y} - \frac{\partial 
\mathcal{H}_y}{\partial z}.
\end{align}
The two remaining velocity components are computed from the definitions of $f$ 
and $g$. The biharmonic equation (\ref{eq:biharmonic}) is solved with boundary 
conditions are $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann conditions follow 
from the continuity equation. The boundary conditions for Eq. (\ref{eq:g}) are 
$g(\pm 1) = 0$. The variational formulation becomes: Find ${u} \in 
\N{V}_N$, ${g} \in \D{V}_N$ and ${f} \in \D{V}_N$ such that
\begin{align}
	\frac{\partial }{\partial t} \left< \nabla^2 u, \N{\psi}\right>_w &= 
	\left<h_u, \N{\psi} \right>_w + \nu \left<\nabla^4u, \N{\psi}\right>_w 
	&\forall \N{\psi} \in \N{V}_N, \label{eq:u1} \\
	\frac{\partial}{\partial t}\left<g, \D{\psi}\right>_w &= \left<h_g, 
	\D{\psi}\right>_w + \nu 
	\left<\nabla^2 g, \D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N ,
	\label{eq:g1} \\
	\left<f, \D{\psi}\right>_w &= \left<\frac{\partial u}{\partial x}, 
	\D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N. \label{eq:f1}
\end{align}
The two remaining velocity components are computed by projection to the 
Dirichlet space $\D{V}_N$: Find 
${v}, {w}$ in $\D{V}_N \times \D{V}_N$ such that
\begin{align}
\left<f, \D{\psi}\right>_w &= \left<\frac{\partial v}{\partial y} + 
\frac{\partial w}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
\D{V}_N, \label{eq:f2} \\
\left<g, \D{\psi}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
\frac{\partial v}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
\D{V}_N, \label{eq:g2}
\end{align}
which simplifies considerably because all the derivatives are in periodic 
directions. Written in spectral space Eqs. (\ref{eq:f2}) and (\ref{eq:g2}) 
become simply
\begin{align}
\hat{f}_{\bm{k}} &= \imath \underline{m}\, \hat{v}_{\bm{k}} + \imath 
\underline{n}\, \hat{w}_{\bm{k}} &\forall \bm{k} \in \D{K}_N^0, \label{eq:f3} \\
\hat{g}_{\bm{k}} &= \imath \underline{m}\, \hat{w}_{\bm{k}} - \imath 
\underline{n}\, \hat{v}_{\bm{k}} & \forall \bm{k} \in \D{K}_N^0, \label{eq:g3}
\end{align}
where $\D{K}_N^0$ is used to denote that these equations can be solved for all 
wavenumbers except $m=n=0$. For $m=n=0$ we solve instead the 
corresponding momentum equations in $y$ and $z$ directions:
\begin{align}
\frac{\partial }{\partial t} \D{B}_{lq}\hat{v}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) + \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) 
& \forall\, l \in [0, 1, 
\ldots, N_x-2], \label{eq:v0}\\
\frac{\partial }{\partial t} \D{B}_{lq}\hat{w}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_z(q, 0, 0, t) + \nu \D{A}_{lq} \hat{w}(q, 0, 0, t) 
& \forall\, l \in [0, 1, \ldots, N_x-2].\label{eq:w0}
\end{align}
